---
title: "Stat 450"
author: 'R Assignment 1: Sam Dokkebakken'
date: "Due Thursday, September 1 by 5pm on your GitHub repositories"
output: word_document
---
Good overall; 18/18.  See comments below.
**Instructions:** The entirety of this assignment must be submitted as an R Markdown file (.Rmd) on your GitHub repository.  Use the .Rmd note handout files and the R Markdown cheat sheet as guidelines.  You are encouraged to save this HW1.Rmd file and fill in the questions with your answers, then submit.  **I should be able to knit your .Rmd file and compile your code myself, so make sure you do some bug checks before submitting!** (I.e., knit the document yourself a couple times and search for errors.)

Consider Example 2 in the notes.  2 dice are rolled, one red and one white.  Let $Y$ be the random variable that denotes the maximum value of the two rolls.  We will use simulation to find the mean and variance of $Y$, and then verify that our simulated results match what we would expect theoretically.

**Theoretical section**

1. (3pts) Define the pmf, find $\mu = E(Y)$, $\sigma^2 = Var(Y)$, and $\sigma = SD(Y)$.  Show all your work.
```{r,eval=TRUE}
Max=1:6
Probability=c(1/36,3/36,5/36,7/36,9/36,11/36)
pmf=data.frame(x=Max,y=Probability)

mean1=1*(1/36)+2*3/36+3*5/36+4*7/36+5*9/36+6*11/36
mean1
var1=(((1-mean1)^2)*1/36)+(((2-mean1)^2)*3/36)+(((3-mean1)^2)*5/36)+(((4-mean1)^2)*7/36)+(((5-mean1)^2)*9/36)+(((6-mean1)^2)*11/36)
var1
sigma1=((var1)^(1/2))
sigma1
```

2. (2pts)  Use `ggplot()` to plot the pmf; see Handout 1 notes for an example.
```{r,eval=TRUE}
library(ggplot2)
y=1:6
py=c(1/36,3/36,5/36,7/36,9/36,11/36)
dd= data.frame(y=y,probs=py)
ggplot(aes(x=y,y=probs),data=dd)+geom_bar(stat='identity')+ylab('P(Y=y)')+ggtitle('pmf of Y')
```

3. (2pts)  Consider the random variable $Z=2Y+1$.  What is $E(Z)$ and $Var(Z)$?  Show all work. 
```{r,eval=TRUE}
expected = 2*mean1+1
expected
var2 = (((3-expected)^2)*1/36)+(((5-expected)^2)*3/36)+(((7-expected)^2)*5/36)+(((9-expected)^2)*7/36)+(((11-expected)^2)*9/36)+(((13-expected)^2)*11/36)
var2

```
OK; or could just take 4*Var(Y)
**Simulation section**  


Write a function called `one.Y` that simulates rolling two dice and returns the maximum roll.  Try the function out a few times and include the results of these test-runs in your R Markdown output. I have written some code below to get you started; each line of "pseudo-code" should be repaced with actual code:

<!--NOTE: You should probably change the below code to eval=TRUE when you submit, else your function will not be compiled and your pdf/word/HTML file will not reproduce any output!  -->

```{r,eval=TRUE}
library(ggplot2)
one.roll = function(){
  sample.space=1:6
  first.roll=sample(sample.space,1,replace=TRUE)
  second.roll=sample(sample.space,1,replace=TRUE)
  max.roll=max(first.roll,second.roll)
  return(max.roll)
}
one.roll()
many.rolls = replicate(1000,one.roll())
df1=data.frame(x=as.factor(many.rolls))
ggplot(aes(x=as.factor(many.rolls)),data=df1)+geom_bar(aes(y=(..count..)/(sum(..count..))))+ylab('Observed Proportion')+xlab('Max')+ggtitle('Simulated pmf')
```

Each of the following can be answered with 1-2 lines of R code (and corresponding output, of course)

4. (2pts) Use `replicate()` to simulate the results of 1000 pairs of rolls.  These are 1000 realizations of the random variable $Y$. Save the 1000 realizations in an object called `many.Y`.
```{r,eval=FALSE}
many.rolls = replicate(1000,one.roll())
```

5.  (2pts) Use `ggplot()` to create the empirical (i.e., observed) pmf of your simulation.  See Handout 1 for example R code.  How does it compare with your theoretical pmf?
```{r,eval=TRUE}
library(ggplot2)
df1=data.frame(x=as.factor(many.rolls))
ggplot(aes(x=as.factor(many.rolls)),data=df1)+geom_bar(aes(y=(..count..)/(sum(..count..))))+ylab('Observed Proportion')+xlab('Max')+ggtitle('Simulated pmf')
```

6. (1pt)  What is the mean of the 1000 realizations?
```{r,eval=TRUE}
mean(many.rolls)
```

7. (1pt)  What is the variance of the 1000 realizations?
```{r,eval=TRUE}
var(many.rolls)
```

8.  (1pt) What is the standard deviation of the 1000 realizations?
```{r,eval=TRUE}
sd(many.rolls)
```

9. (1pt)  Create a new object called `many.Z` that creates 1000 realizations of $Z$.
```{r,eval=TRUE}
z.roll=function(){
  sample.space=c(3,5,7,9,11,13)
  a.roll = sample(sample.space,1,replace=TRUE)
  b.roll = sample(sample.space,1,replace=TRUE)
  c.roll = max(a.roll,b.roll)
  return(c.roll)
}
z.roll()
alotta.rolls=replicate(1000,z.roll())
```
OK; could just take Z <- 2*Y+1 and calculate mean, variance from there.
10. (1pt)  What is the mean of $Z$?
```{r,eval=TRUE}
mean(alotta.rolls)
```

11. (1pt)  What is the variance of $Z$?
```{r,eval=TRUE}
var(alotta.rolls)
```

12. (1pt) Note that your simulated results should be similar to the theoretical quantities; if they aren't, re-check your R code!  What is the reason for any differences?
They are very similar, however there is a slight difference. For the expected values of mu and variance, everyone should get the same answer. However, when we run a 1000 replications, each time you do run 1000 replications, we will get slightly different answers.

